Brief Article

### 2012-02-14 power.tex
corrected typos
some updates

Now assume that the same lineup plot is evaluated by $K$ independent observers. Wlog let the last plot be the data plot and plots 1 through $m-1$ the null plots. Let $n_i$ be the number of people who picked plot $i$ as the plot with the strongest expression of non-random structure, then $\sum_{k=1}^{m} n_k = K$

For the distribution of $n_m$, the number of times out of $K$ that observers picked the last plot to show the most non-random structure, we get
\[
n_m\sim \left \{ 
\begin{array}{cl}
B_{K, (1-E[p_B])^{m-1} & \text { under } H_1\\
B_{K, 1/m} & \text { under } H_0
\end{array}
\right .
\]
Figure \ref{power} gives an overview of the probability of picking the dataplot for lineups of different size (as $m$ increases we have an increased probability to observe a more highly structured nullplot by chance).
Even under $H_1$ we expect some observers to pick a nullplot with a probability that depends on the strength of the signal in our dataplot, i.e. reversely, we should be able to make use of the number of people who do not pick the dataplot in a lineup, to infer the strength of the signal of the dataplot.

\begin{figure}[htbp] %  figure placement: here, top, bottom, or page
   \centering
   \includegraphics[width=2in]{images/powerplot.pdf} 
   \caption{Comparison of probabilities to pick the data plot for different size lineups $m$ and different values for data plot strength $p_B$. The grey line shows probability of picking the data plot under $H_0$. We see that for a $p$ value $p_B$ of about 0.15 for the dataplot the signal in the plot is so weak, that we cannot distinguish the dataplot from nullplots in a lineup of size $m=20$. (Mahbub, is that what we see in practice?)}
   \label{power}
\end{figure}

\end{document}

>>>>

Now assume that the same lineup plot is evaluated by $K$ independent observers. Wlog let the last plot be the data plot and plots 1 through $m-1$ the null plots. Let $n_i$ be the number of people who picked plot $i$ as the plot with the strongest expression of non-random structure, then $\sum_{k=1}^{m} n_k = K$, and $(n_1, n_2, ..., n_m) \sim$ Mult$_{\pi_1, ..., \pi_m}$ with $\sum_i \pi_i = 1$. We can estimate $\pi_i$ as $\widehat{\pi_i} = n_i/K$.

For the distribution of $n_m$, the number of times out of $K$ that observers picked the last plot to show the most non-random structure, we get
\[
n_m\sim \left \{ 
\begin{array}{cl}
B_{K, (1-E[p_B])^{m-1}} & \text { under } H_1\\
B_{K, 1/m} & \text { under } H_0
\end{array}
\right .
\]
Figure \ref{power} gives an overview of the probability of picking the dataplot for lineups of different size (as $m$ increases we have an increased probability to observe a more highly structured nullplot by chance).
Even under $H_1$ we expect some observers to pick a nullplot with a probability that depends on the strength of the signal in our dataplot, i.e. reversely, we should be able to make use of the number of people who do not pick the dataplot in a lineup, to infer the strength of the signal of the dataplot.

\begin{figure}[htbp] %  figure placement: here, top, bottom, or page
   \centering
   \includegraphics[width=2in]{images/powerplot.pdf} 
   \caption{Comparison of probabilities to pick the data plot for different size lineups $m$ and different values for data plot strength $p_B$. The grey line shows probability of picking the data plot under $H_0$. We see that for a $p$ value $p_B$ of about 0.15 for the dataplot the signal in the plot is so weak, that we cannot distinguish the dataplot from nullplots in a lineup of size $m=20$. (Mahbub, is that what we see in practice?)}
   \label{power}
\end{figure}
Under $H_1$ we therefore get an estimate for the strength of the dataplot signal $p_B$ as
\[
\widehat{E[p_B]} = 1 - \left( \frac{n_m}{K} \right)^{1/(m-1)}.
\]
Similarly, we get an estimate for $E[X]$, the average number of null plots with a stronger signal strength than the data plot as the number of nullplots in the lineup with more than $n_m$ picks in their favor, which leads us to another estimate of $E[p_B]$ as 
\[
\widehat{E[p_B]}  = 1 - 1/(m-1) \cdot \sum_i I (n_i > n_m),
\]
where $I(x) = 1$ for $x$ is true and $I(x) = 0$ if $x$ is false.
\end{document}
